---
title: "Introduction to Regression in R"
output:
  pdf_document: default
---

# Simple Linear Regression

- Response variable (a.k.a. dependent variable): The variable that we want to predict

- Explanatory variables (a.k.a. independent variables): The variables that explain how the response variable will change.

```{r, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)

auto_insurance <- read_csv('https://raw.githubusercontent.com/truonghm/DataCamp/main/Statistician_with_R/3_Introduction_to_Regression_in_R/sweden_auto_insurance.csv', 
         col_types=cols(
             X = col_double(),
             Y = col_double()
             )
         )
auto_insurance %>%
    summarise(cor(X, Y))

# lm means linear model
ggplot(data=auto_insurance, mapping=aes(x=X, y=Y)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE)
```

## Numerical explanatory variables 

### Taiwan real estate dataset

The Taiwan real estate dataset contains 4 variables:

| dist_to_mrt_station_m | Distance to nearest MRT metro station, in meters.                   |
|-----------------------|---------------------------------------------------------------------|
| n_convenience         | No. of convenience stores in walking distance.                      |
| house_age_years       | The age of the house, in years, in 3 groups.                        |
| price_twd_msq         | House price per unit area, in New Taiwan dollars per meter squared. |

```{r}
taiwan_real_estate <- read_csv('https://raw.githubusercontent.com/truonghm/DataCamp/main/Statistician_with_R/3_Introduction_to_Regression_in_R/taiwan_real_estate.csv', 
         col_types = cols(
             dist_to_mrt_m = col_double(),
             n_convenience = col_double(),
             house_age_years = col_character(),
             price_twd_msq = col_double()
             )
         )
# The number of convenience stores contains integer data, causing points to overlap. 
# To solve this, we will make the points transparent.
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method='lm', se=FALSE)
```

### Fitting a linear regression

from the code below, we can see that the data fit to the following linear equation: $Y = 8.2242 + 0.7981  X$.

```{r}
# Run a linear regression of price_twd_msq vs. n_convenience
lm(price_twd_msq ~ n_convenience, data = taiwan_real_estate)
```

From the coefficients above, we can interpret that:

- On average, a house with zero convenience stores nearby had a price of $8.2242$ TWD per square meter.

- If you increase the number of nearby convenience stores by one, then the expected increase in house price is $0.7981$ TWD per square meter.

## Categorical explanatory variables
