---
title: "Introduction to Regression in R"
output:
  pdf_document: default
---

# Simple Linear Regression

- Response variable (a.k.a. dependent variable): The variable that we want to predict

- Explanatory variables (a.k.a. independent variables): The variables that explain how the response variable will change.

```{r, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)

auto_insurance <- read_csv('https://raw.githubusercontent.com/truonghm/DataCamp/main/Statistician_with_R/3_Introduction_to_Regression_in_R/sweden_auto_insurance.csv', 
         col_types=cols(
             X = col_double(),
             Y = col_double()
             )
         )
auto_insurance %>%
    summarise(cor(X, Y))

# lm means linear model
ggplot(data=auto_insurance, mapping=aes(x=X, y=Y)) +
    geom_point() +
    geom_smooth(method='lm', se=FALSE)
```

## Numerical explanatory variables 

### Taiwan real estate dataset

The Taiwan real estate dataset contains 4 variables:

| dist_to_mrt_station_m | Distance to nearest MRT metro station, in meters.                   |
|-----------------------|---------------------------------------------------------------------|
| n_convenience         | No. of convenience stores in walking distance.                      |
| house_age_years       | The age of the house, in years, in 3 groups.                        |
| price_twd_msq         | House price per unit area, in New Taiwan dollars per meter squared. |

```{r}
taiwan_real_estate <- read_csv('https://raw.githubusercontent.com/truonghm/DataCamp/main/Statistician_with_R/3_Introduction_to_Regression_in_R/taiwan_real_estate.csv', 
         col_types = cols(
             dist_to_mrt_m = col_double(),
             n_convenience = col_double(),
             house_age_years = col_character(),
             price_twd_msq = col_double()
             )
         )
# The number of convenience stores contains integer data, causing points to overlap. 
# To solve this, we will make the points transparent.
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method='lm', se=FALSE)
```

### Fitting a linear regression

from the code below, we can see that the data fit to the following linear equation: $Y = 8.2242 + 0.7981  X$.

```{r}
# Run a linear regression of price_twd_msq vs. n_convenience
lm(price_twd_msq ~ n_convenience, data = taiwan_real_estate)
```

From the coefficients above, we can interpret that:

- On average, a house with zero convenience stores nearby had a price of $8.2242$ TWD per square meter.

- If you increase the number of nearby convenience stores by one, then the expected increase in house price is $0.7981$ TWD per square meter.

## Categorical explanatory variables

The Taiwan real estate dataset has a categorical variable in the form of the age of each house. The ages have been split into 3 groups: 0 to 15 years, 15 to 30 years, and 30 to 45 years.

```{r}
# Using taiwan_real_estate, plot price_twd_msq
ggplot(data=taiwan_real_estate, mapping=aes(x=price_twd_msq)) +
  # Make it a histogram with 10 bins
  geom_histogram(bins=10) +
  # Facet the plot so each house age group gets its own panel
  facet_wrap(vars(house_age_years))
```

### Calculating means by category

A good way to explore categorical variables is to calculate summary statistics such as the mean for each category.

```{r}
summary_stats <- taiwan_real_estate %>% 
  # Group by house age
  group_by(house_age_years) %>% 
  # Summarize to calculate the mean house price/area
  summarise(mean_by_group = mean(price_twd_msq))

# See the result
summary_stats
```

### Linear regression with a categorical explanatory variable

Linear regressions also work with categorical explanatory variables. In this case, the code to run the model is the same, but the coefficients returned by the model are different.

In the code below, the intercept is actually the mean price of 0 to 15, and the other 2 coefficients are relative difference to from the intercept.

```{r}
# Run a linear regression of price_twd_msq vs. house_age_years
mdl_price_vs_age <- lm(price_twd_msq ~ house_age_years, data=taiwan_real_estate)

# See the result
mdl_price_vs_age
```

By changing the model to append `+ 0`, we have the absolute mean price of each age group.

```{r}
# Update the model formula to remove the intercept
mdl_price_vs_age_no_intercept <- lm(
  price_twd_msq ~ house_age_years + 0, 
  data = taiwan_real_estate
)

# See the result
mdl_price_vs_age_no_intercept
```


# Predictions and model objects

Making prediction using linear regression:

```{r}
explanatory_data <- tibble(
  n_convenience = 0:10
)
prediction_data <- explanatory_data %>%
    mutate(price_twd_msq  = predict(mdl_price_vs_conv, explanatory_data)
    )
prediction_data
```

Plotting the prediction data against the original data

```{r}
# Add to the plot
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  # Add a point layer of prediction data, colored yellow
  geom_point(data=prediction_data, aes(n_convenience, price_twd_msq), color='yellow')
```
